---
title: "p8105_hw1_lg3450"
author: "Luyun Ge"
date: "2025-09-20"
output: github_document
---

R Markdown file by Luyun Ge for P8105 Homework1.

We start by loading required libraries.
```{r}
library(moderndive)
library(ggplot2)
library(tidyverse)
```

# Problem1
We load the required dataset for problem1.
```{r}

data("early_january_weather")
```
Then we list all variables in this dataset, calculate the size of this dataset and the mean temperature.
```{r}
names(early_january_weather) #listed all variables in dataset
nrow(early_january_weather) #number of rows
ncol(early_january_weather) #number of columns
mean(early_january_weather$temp) #mean temperature
```
The `early_january_weather dataset` has 15 variables: `origin`, `year`, `month`, `day`, `hour`, `temp`, `dewp`, `humid`, `wind_dir`, `wind_speed`, `wind_gust`, `precip`, `pressure`, `visib`, `time_hour`. 
The dataset has 358 rows and 15 columns. 
The mean temperarture of early January is 39.58212 Fahrenheit.

Then we plot the scatterplot of temperature vs time.
```{r}
p1_plot <- ggplot(early_january_weather, aes(x = time_hour, y = temp, color = humid)) +
  geom_point() +
  labs(title = "Temperature vs Time (in hours)",
       x = "Time (in hours)",
       y = "Temperature",
       color = "Humidity")
p1_plot
ggsave("temp_vs_time_hour.png", plot = p1_plot)
```
The scatterplot shows that temperature fluctuates across the hours of the day over early January, generally ranging from about 25 to 50 degrees Fahrenheit.Higher humidity values (lighter blue points) occur more often at higher temperatures, while lower humidity values (darker blue points) occur at lower temperatures, indicating a positive relationship between temperature and humidity over time.

# Problem2
To solve the problem, we first create the data frame based on requirements.
```{r}
p2_df <- data.frame(
  random_Nsample = rnorm(10),
  greater_than_zero = rnorm(10) > 0,
  character = letters[1:10],
  factor = factor(rep(c("A", "B", "C"), length.out = 10))
)
p2_df
```
Then calculate the mean of each variables in the data frame.
```{r}
mean(pull(p2_df,random_Nsample))
mean(pull(p2_df,greater_than_zero))
mean(pull(p2_df,character))
mean(pull(p2_df,factor))  
```
The mean could be calculated for `random_Nsample` because the mean function works for numeric values. 
The mean could also be calculated for `greater_than_zero` because R converts TRUE to 1 and FALSE to 0.
We failed to calculate the mean for `character` and `factor` because characters and categorical factors cannot be averaged.
Now we convert all variables to numeric variables.
```{r}
as.numeric(p2_df$greater_than_zero)
as.numeric(p2_df$character)
as.numeric(p2_df$factor)
```
We are able to convert the `greater_than_zero` and `factor` variables to numeric, but conversion fails for the `character` variable. The logical variable `greater_than_zero` converts to 0 and 1, so calculating its mean is meaningful. `factor` converts to underlying integer codes; while the mean produces a number, it does not meaningfully represent the categories.






